---
title: "Regression and Analysis of Variance"
author: "Dataset 3 (Fish Market)."
date: "29/11/2020"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

## Fish Market Dataset for Regression
#### Data Description.

El conjunto de datos del mercado de pescado contiene información sobre especies de peces comunes en las ventas del mercado. El conjunto de datos incluye las especies de peces, peso, longitud, altura y ancho.

Este conjunto de datos es un registro de 7 especies de peces diferentes comunes en las ventas del mercado de pescado. Con este conjunto de datos, se puede realizar un modelo predictivo utilizando datos fáciles de usar y estimar el peso de los peces.


```{r}
data = read.csv('Fish.csv')
head(data)
```

Meanings of the Columns:

Species: Species name of fish
Weight: Weight of fish in gram
Length1: Vertical length in cm
Length2: Diagonal length in cm
Length3: Cross length in cm
Height: Height in cm
Width: Diagonal width in cm
Our dependent variable is 'Weight'. Independent variables are 'species', different lengths, 'height' and 'width'.

I will use independent variables (measurements of the fish) to estimate dependent variable (weight of the fish).
```{r data}
summary(data)
```
```{r}
str(data)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
